---
title: "What is a reprex?"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{What is a reprex?}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
library(reprex)
```

## Writing a good reprex

Where and why are they used?

* A Stack Overflow question that includes a proper reprex is [much more likely to get answered](http://stackoverflow.com/help/no-one-answers), by the most knowledgeable (and therefore busy!) people.

* A [GitHub issue](https://guides.github.com/features/issues/) that includes a proper reprex is more likely to achieve your goal: getting a bug fixed or getting a new feature, in a finite amount of time.

What are the main requirements?

* Use the smallest, simplest, most [built-in data](https://stat.ethz.ch/R-manual/R-patched/library/datasets/html/00Index.html) possible.
    - Think: `iris` or `mtcars`. Bore me.
    - If you must make some objects, minimize their size and complexity.
    - Get just a bit of something with `head()` or by indexing with the result of `sample()`. If anything is random, consider using `set.seed()` to make it repeatable.
    - `dput()` is a good way to get the code to create an object you have lying around. Copy and paste the *result* of this into your reprex.
    - Look at official examples and try to write in that style. Consider adapting one.

* Include commands on a strict "need to run" basis.
    - Ruthlessly strip out anything unrelated to the specific matter at hand.
    - Include every single command that is required, e.g. loading specific packages via `library(foo)`.

* Consider including so-called "session info", i.e. your OS and versions of R and add-on packages, if it's conceivable that it matters. Use `reprex(..., si = TRUE)` for this.

* Whitespace rationing is not in effect. Use good [coding style](http://adv-r.had.co.nz/Style.html).

* Pack it in, pack it out, and don't take liberties with other people's computers. You are asking people to run this code!
    - If you change options, store original values at the start, do your thing, then restore them: `opar <- par(pch = 19) <blah blah blah> par(opar)`.
    - If you create files, delete them when you're done: `write(x, "foo.txt") <blah blah blah> file.remove("foo.txt")`.
    - Don't delete files or objects that you didn't create in the first place.
    - Don't mask built-in functions, i.e. don't define a new function named `c`.
    - Take advantage of R's built-in ability to create temporary files and directories. Read up on [`tempfile()` and `tempdir()`](https://stat.ethz.ch/R-manual/R-devel/library/base/html/tempfile.html).
    - Don't start with `setwd("C:\Users\jenny\path\that\only\I\have")`, because it won't work on anyone else's computer.
    - Don't start with `rm(list = ls())`, because it is anti-social to clobber other people's workspaces.
  
But won't that take time and effort?

* Yes, yes it will!

* 80% of the time you will solve your own problem in the course of writing an excellent reprex. YMMV.

* The remaining 20% of the time, you will create a reprex that is more likely to elicit the desired behavior in others.

Get more concrete details here:

* [How to make a great R reproducible example?](http://stackoverflow.com/questions/5963269/how-to-make-a-great-r-reproducible-example/16532098) thread on StackOverflow

* [How to write a reproducible example](http://adv-r.had.co.nz/Reproducibility.html) from Hadley Wickham's [Advanced R book](http://adv-r.had.co.nz)

## Package philosophy

The reprex code:

* Must run and, therefore, should be run **by the person posting**. No faking it.

* Should be easy for others to digest, so **they don't necessarily have to run it**. You are encouraged to include selected bits of output. :scream:

* Should be easy for others to copy + paste + run, **iff they so choose**. Don't let inclusion of output break executability.

Accomplished like so:

* use `rmarkdown::render()` or, under the hood, `knitr::spin()` to run the code and capture output that would display in R console

* use chunk option `comment = "#>"` to include the output while retaining executability

## Advanced features

### Embedded prose

Sometimes you want to mingle rendered code and prose. Put the embedded prose in as roxygen comments, i.e. comment lines that start with `#'`. This reprex code:

```{r include = FALSE}
demo_code <- c(
  "## a regular comment",
  "x <- 1:100",
  "#' Here is some embedded prose, as a roxygen comment.",
  "mean(x)"
)
md_content <-
  reprex(input = demo_code, outfile = "README-tmp_gh", show = FALSE)
```

```{r asis = TRUE, echo = FALSE, comment = NA}
cat(demo_code, sep = "\n")
```

renders to this this result:

```{r child = "README-tmp_gh_reprex.md"}
```

```{r include = FALSE}
leftovers <- list.files(pattern = "README-tmp", ignore.case = TRUE)
if (length(leftovers)) file.remove(leftovers)
```


## Other work

If I had known about [`formatR::tidy_eval()`](http://yihui.name/formatR/), I probably would never had made reprex! But alas I did not. AFAICT here are the main differences:

  * `reprex()` accepts an expression as primary input, in addition to code on the clipboard, in a character vector, or in a file.
  * `reprex()` runs the reprex in a separate R process, via [callr](https://cran.r-project.org/package=callr). `tidy_eval()` uses the existing R process and offers an `envir` argument.
  * `reprex()` writes the code to a `.R` file and calls `rmarkdown::render()`. `tidy_eval()` runs the code line-by-line via `capture.output(eval(..., envir = envir))`.
  * `reprex()` uploads figures to imgur and inserts the necessary link.
