---
title: "How to suppress package startup messages"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{How to suppress package startup messages}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

When making a reprex, you might need to attach packages that are producing some messages and warnings upon attachment. The goal of this article is to suppress these, without suppressing all messages and warnings of the reprex, as they might be useful for debugging. This article aims at answering the issue [#187](https://github.com/tidyverse/reprex/issues/187). 

We review below the different possibilities to silence messages and warnings when attaching a package. We use dplyr as an example.         

## Short version 

Messages and warnings can be suppressed by using the `warn.conflicts` and `quietly` arguments in `library()` or by using the `suppressPackageStartupMessages()` function.   

```{r, eval=FALSE}
library(dplyr, warn.conflicts = FALSE)
library(dplyr, quietly = TRUE)
suppressPackageStartupMessages(library(dplyr))
```

## Attaching a package

Attaching dplyr with `library(dplyr)` produces two warnings about conflicts. We would like to silence them.       

```{r}
# Attaching dplyr
library(dplyr)
```
    

## Attaching a package without warnings

To suppress the warnings about conflicts, you can set the `warn.conflicts` argument to `FALSE` in `library()`. As written in the documentation of the `library()` function, a conflict is a function masking a function, or a non-function masking a non-function. Note that this also silences the line `Attaching package: ‘dplyr’`.     

```{r}
# Detaching dplyr (attached in the chunk above)
detach("package:dplyr", unload=TRUE)
# Attaching dplyr
library(dplyr, warn.conflicts = FALSE)
```


## Attaching a package without messages

To suppress some of the startup messages, you can also set the `quietly` argument in the `library()` function to `TRUE`. This does not suppress the warnings about conflicts. Also note that here the line `Attaching package: ‘dplyr’` is not affected by `quietly`.

```{r}
# Detaching dplyr (attached in the chunk above)
detach("package:dplyr", unload=TRUE)
# Attaching dplyr
library(dplyr, quietly = TRUE)
```


## Attaching a package silently with the function suppressPackageStartupMessages()

`suppressPackageStartupMessages()` ignores diagnostic messages when attaching a package. It also ignores warnings about conflicts. Here it suppresses all messages and warnings. 
```{r}
# Detaching dplyr (attached in the chunk above)
detach("package:dplyr", unload=TRUE)
# Attaching dplyr
suppressPackageStartupMessages(library(dplyr))
```


## Attaching a package silently within a chunk

Another way to silence the warnings and messages is to set the chunk options `message` and `warning` to `FALSE`. To prevent hiding useful warnings and messages for debugging a reprex, separate the chunk attaching the packages from the rest of the reprex. 

Within an Rmarkdown document that you `knit`:

````r
`r ''````{r chunk1, message=FALSE, warning=FALSE}  
library(dplyr)  
```
Some text
`r ''````{r chunk2, message=TRUE, warning=TRUE}     
# reprex code goes here ... 
```
````

or within an R script that you `spin`: 

```{r}
#+ message = FALSE, warning = FALSE
library(dplyr)

#+ reprex-body
# reprex code goes here ...
```


## Attaching the tidyverse package silently

In the function `reprex()` from reprex, the argument `tidyverse.quiet` (set to `TRUE` by default) suppresses the startup messages for the tidyverse package.

```{r, eval=FALSE}
# Attaching reprex
library(reprex)

# Creating a reprex in which tidyverse attachment is not quiet
ret <- reprex({
library(tidyverse)
a <- 1:5
tibble(a, a * 2)
}, 
tidyverse_quiet = FALSE)
```

Alternatively, you can open the .Rprofile configuration file with the function `edit_r_profile()` from the package `usethis` and add `options(tidyverse.quiet = TRUE)` to the file. This will affect all calls to `reprex()`.
```{r, eval=FALSE}
# Attaching usethis
library(usethis)
# Opening .Rprofile configuration file
edit_r_profile()
```


