---
title: "Running an external process on Windows"
output: github_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  error = TRUE
)
```

Figuring out how to write RTF to the Windows clipboard made me think more about the functions `system()`, `system2()`, and (Windows only) `shell()`.

Here's the highlight call used in `reprex_rtf()`, which converts such a `.R` file into RTF:

```{sh eval = FALSE}
highlight -i foo_reprex.R --out-format rtf --no-trailing-nl --encoding=UTF-8 --style darkbone --font 'Courier Regular' --font-size 50 -o gamma_reprex.rtf
```

There are many ways to invoke this from R:

* `system()`: Docs say "not recommended for new code".
* `system2()`: Officially, the "right" thing to use, at least on *nix.
* `shell()`: Exists only on Windows. Explicitly runs `cmd` under a shell, which
  is not the case for `system()` or `system2()` on Windows.
* `processx::run()`

`system()` and `system2()` only differ with respect to how the ultimately-run command is constructed.
`system2()` is recommended for new code and takes care of some fiddly details, e.g. quoting.
However, due to quoting nightmares, there are calls that can only be formed by `system()` and some that can't be formed at all (which is one of the reasons processx exists).
The main thing to know is that on Windows `command` is executed directly, not within a shell, such as `cmd.exe` or `powershell.exe`.

This is easiest to see in an example.
`dir` is a command built into `cmd.exe`.
It will work inside `shell()` but not within `system()` or `system2()`.

```{r}
shell("dir", intern = TRUE)
print(system("dir"))
print(system2("dir"))
```

How did this come up?
Depending on how you have installed highlight, the `system()` and `shell()` calls below might produce different results.
Specifically, the `system()` call is reported to fail here if highlight was installed via scoop, which uses a shim-based method of putting things it installs on the `PATH`.
This may additionally have something to do with Windows short paths (?).

```{r}
system("highlight --version")
shell("highlight --version")
Sys.which("highlight")
```

In any case, my current hypothesis is that I should invoke highlight via `shell()` on Windows, to be more resilient to the different ways users may have installed highlight.

Let's prove I can invoke highlight in several ways by implementing `reprex_rtf()` "by hand".
First, we render a reprex to the "r" venue and explicitly write a local output file.

```{r}
library(reprex)

reprex_r(sample(LETTERS, 5), outfile = here::here("internal/gamma"))
```

The output of this `gamma_reprex_rendered.R` is the input for the next step, which is to form RTF.

More objects and helpers we'll use:

```{r}
r_file <- here::here("internal/gamma_reprex_rendered.R")
rtf_file <- here::here("internal/gamma_reprex_rendered.rtf")
```

### `system()`

```{r}
(cmd <- glue::glue("
   highlight -i {r_file} --out-format=rtf --no-trailing-nl --encoding=UTF-8 \\
   -o {rtf_file}"))

system(cmd)

readLines(rtf_file)
```

### `system2()`

```{r}
unlink(rtf_file)

(args <- c(
  glue::glue("-i {r_file}"),
  "--out-format=rtf", "--no-trailing-nl", "--encoding=UTF-8",
  glue::glue("-o {rtf_file}")
))

print(system2("highlight", args))

readLines(rtf_file)
```

### `shell()`

```{r}
unlink(rtf_file)

(cmd <- glue::glue("
   highlight -i {r_file} --out-format=rtf --no-trailing-nl --encoding=UTF-8 \\
   -o {rtf_file}"))

print(shell(cmd))

readLines(rtf_file)

unlink(rtf_file)

print(shell(cmd, shell = "cmd"))

readLines(rtf_file)

unlink(rtf_file)

print(shell(cmd, shell = "powershell"))

readLines(rtf_file)
```

### Clean up

```{r}
unlink(list.files(here::here("internal"), pattern = "gamma_reprex", full.names = TRUE))
```
