---
title: "Windows clipboard data persistence"
output: github_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  error = TRUE
)
```

Document what I learned about data persistence while writing to the Windows clipboard

First, write a function to expose RTF on the clipboard from R, then call it to establish baseline:

```{r}
read_clipboard_rtf <- function() {
  stopifnot(.Platform$OS.type == "windows")
  system(
    'powershell -Command "Get-Clipboard -TextFormatType Rtf"',
    # `intern = TRUE` is a concession to knitr and making sure we see what
    # `Get-Clipboard` returns in our rendered result
    # it would not be necessary if you execute this interactively
    intern = TRUE)
}
read_clipboard_rtf()
```

The approach I ultimately used in reprex based on the `SetText` method:

```{r}
rtf <- r"({\rtf1\ansi\deff0 {\fonttbl {\f0 Times New Roman;}} \f0\fs60 Hello, World!})"
cmd <- glue::glue('
  powershell -Command "\\
  Add-Type -AssemblyName System.Windows.Forms | Out-Null;\\
  [Windows.Forms.Clipboard]::SetText(\\
  \'{rtf}\',\\
  [Windows.Forms.TextDataFormat]::Rtf\\
  )"')
system(cmd)
read_clipboard_rtf()
```

The more complicated approach based on `SetDataObject`, before I appreciated the persistence problem:

```{r}
rtf <- r"({\rtf1\ansi\deff0 {\fonttbl {\f0 Times New Roman;}} \f0\fs60 Hello, World!})"
cmd <- glue::glue('
  powershell -Command "\\
  Add-Type -AssemblyName System.Windows.Forms | Out-Null;\\
  $data = New-Object Windows.Forms.DataObject
  $data.SetData([Windows.Forms.DataFormats]::Rtf, \'{rtf}\')
  [Windows.Forms.Clipboard]::SetDataObject($data)"')
system(cmd)
read_clipboard_rtf()

```

Adding the all important `$true` to the `SetDataObject` call:

```{r}
rtf <- r"({\rtf1\ansi\deff0 {\fonttbl {\f0 Times New Roman;}} \f0\fs60 Hello, World!})"
cmd <- glue::glue('
  powershell -Command "\\
  Add-Type -AssemblyName System.Windows.Forms | Out-Null;\\
  $data = New-Object Windows.Forms.DataObject
  $data.SetData([Windows.Forms.DataFormats]::Rtf, \'{rtf}\')
  [Windows.Forms.Clipboard]::SetDataObject($data, $true)"')
system(cmd)
read_clipboard_rtf()

```

Conclusion: `SetText` write persistent data, whereas by default `SetDataObject` does not (although it can if you use the second argument, `copy`).
